---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

devtools::load_all(".")
```

# rdcomonotonicity

`rdcomonotonicity` implements nonparametric extrapolation in sharp regression
discontinuity designs using the comonotonicity approach of Deaner and Kwon
(2025). Under comonotonicity of the conditional mean potential outcomes, this 
package extrapolates average treatment effects at the frontier to observations 
away from the frontier whenever the relevant conditional mean lies within the 
estimated frontier support.

The package currently provide tools to:

- estimate treated and untreated conditional mean outcome functions using
  multivariate local polynomial regression;
- estimate the mappings \(q_0\) and \(q_1\) betweend conditional mean potential
  outcomes;
- determine where conditional average treatment effects are identified;
- construct multiplier-bootstrap confidence bands; and
- estimate the effects of counterfactual treatment policies.



## Installation

Install the development version from GitHub:

```r
# install.packages("pak")
pak::pak("yubnkm/rdcomonotonicity")
```

Then load the package:

```{r load-package, eval = FALSE}
library(rdcomonotonicity)
```


## Toy Example

This example generates a 2-dimensional sharp RDD in which comonotonicity holds 
by construction. It then estimates the potential outcome mappings, obtains 
multiplier bootstrap confidence bands, display the identified region, and 
estimates the effect of a counterfactual change in the treatment frontier.

### Data generating process

Let \(X=(X_1,X_2)\) be uniformly distributed on the unit square. Treatment is
assigned when

\[
X_2 < 0.7 - 0.4X_1.
\]

The conditional mean treated potential outcome is

\[
g_1(x_1,x_2)=\sin(x_1)+0.3\sin(x_2),
\]

and the conditional mean untreated potential outcome is

\[
g_0(x_1,x_2)=0.8-0.8\cos\{g_1(x_1,x_2)\}.
\]

Therefore, \(g_0(x)=q_0(g_1(x))\), where
\(q_0(y)=0.8-0.8\cos(y)\). The two conditional mean potential outcomes are
comonotonic.


```{r simulate-data}
set.seed(1)

g1_true <- function(x1, x2) sin(x1) + 0.3 * sin(x2)
g0_true <- function(x1, x2) 0.8 - 0.8 * cos(g1_true(x1, x2))

frontier <- function(x1) 0.7 - 0.4 * x1
d_func <- function(x1, x2) as.integer(x2 < frontier(x1))


q0_true <- function(y1) 0.8 - 0.8 * cos(y1)
q1_true <- function(y0) acos(pmax(-1, pmin(1, 1 - y0 / 0.8)))

simulate_dgp <- function(n = 500, noise_sd = 0.02) {
  X <- matrix(runif(2 * n), nrow = n, ncol = 2)
  colnames(X) <- c("X1", "X2")

  mu1 <- g1_true(X[, 1], X[, 2])
  mu0 <- g0_true(X[, 1], X[, 2])
  D <- d_func(X[, 1], X[, 2])

  factual_mean <- ifelse(D == 1, mu1, mu0)
  Y <- factual_mean + rnorm(n, mean = 0, sd = noise_sd)

  list(
    Y = Y,
    X = X,
    D = D,
    mu0 = mu0,
    mu1 = mu1,
    factual_mean = factual_mean
  )
}

dat <- simulate_dgp(n = 1000, noise_sd = 0.02)
```

### Estimate the extrapolation model

`RDD_extrapolate_CV_band()` estimates the factual conditional mean outcome
separately in the treated and untreated regions. It then uses observations near
the treatment frontier to estimate:

- \(q_0\), which maps a treated conditional mean into the corresponding
  untreated conditional mean; and
- \(q_1\), which maps an untreated conditional mean into the corresponding
  treated conditional mean.

When more than two candidate bandwidths are supplied, the function selects
first-stage bandwidths by cross-validation.

```{r estimate-model, cache = TRUE}
result <- RDD_extrapolate_CV_band(
  Y = dat$Y,
  X = dat$X,
  D = dat$D,
  kernel = "gaussian",
  bands = seq(0.2, 0.6, length.out = 5),
  num_folds = 5,
  order = 1
)

```

The returned support indicator `result$S` equals one when the observation's
factual conditional mean lies within the estimated domain of the relevant
counterfactual mapping. These are the observations for which the package
extrapolates both conditional mean potential outcomes.

### Multiplier bootstrap

`RDD_extrapolate_bootstrap()` applies independent standard exponential
multipliers to the observation weights and re-estimates the model using the
bandwidths selected in the original fit.

The following example uses 30 bootstrap draws and two parallel workers.
Increasing `bootstrap_iter` generally produces more stable confidence bands but
increases computation time.

```{r bootstrap, cache = TRUE}
set.seed(1)

bootstrap_result <- RDD_extrapolate_bootstrap(
  result = result,
  Y = dat$Y,
  X = dat$X,
  D = dat$D,
  kernel = "gaussian",
  num_folds = 5,
  order = 1,
  points = 100L,
  bootstrap_iter = 30L,
  parallel = TRUE,
  n_cores = 2L
)
```

### Estimated potential-outcome mappings

`plot_q_results()` returns three `ggplot` objects:

1. the estimated \(q_0\) mapping;
2. the estimated \(q_1\) mapping; and
3. a comparison of \(\widehat q_1\) with the inverse relationship implied by
   \(\widehat q_0\).

Under comonotonicity, the population versions of \(q_1\) and \(q_0^{-1}\) should
coincide over their common support. The shaded regions below are 90 percent
pointwise multiplier-bootstrap confidence bands.

```{r construct-q-plots}
q_plots <- plot_q_results(
  result = result,
  points = 100L,
  show_points = FALSE,
  bootstrap_result = bootstrap_result,
  level = 0.90
)
```

#### Mapping from treated to untreated conditional means

```{r plot-q0, out.width="60%", fig.width=6, fig.height=4, fig.align="center"}
q_plots$q0_plot
```

#### Mapping from untreated to treated conditional means

```{r plot-q1, out.width="60%", fig.width=6, fig.height=4, fig.align="center"}
q_plots$q1_plot
```

#### Comparison of \(\widehat q_1\) and \(\widehat q_0^{-1}\)

```{r plot-q-comparison, out.width="60%", fig.width=6, fig.height=4, fig.align="center"}
q_plots$comp_plot
```

### Region with identified conditional average treatment effects

For two-dimensional assignment variables, `plot_idf_region()` displays the
observations for which the estimated conditional average treatment effect is
identified. Crosses indicate `S = 1`, while circles indicate `S = 0`.

The `threshold_function` is used only to label the treated and untreated regions
in the plot. It must return one treatment indicator for each row of `X`.

```{r plot-identified-region}
plot_idf_region(
  X = dat$X,
  result = result,
  threshold_function = function(x) {
    0.4 * x[, 1L] + x[, 2L] < 0.7
  }
)
```


### Counterfactual policy effect

Consider a counterfactual policy that shifts the treatment frontier upward by
0.05:

$$
D^{\mathrm{counterfactual}} = \mathbf{1}\left\{X_2 < 0.7 - 0.4X_1 + 0.05\right\}.
$$

This policy expands treatment to observations immediately above the original
frontier.

```{r define-policy}
delta <- 0.05

counterfactual_policy <- function(X) {
  new_frontier <- frontier(X[, 1L]) + delta
  as.numeric(X[, 2L] < new_frontier)
}
```

`counterfactual_policy_effect()` estimates the mean outcome change generated by
replacing the observed treatment rule with the counterfactual policy. The
reported effect is averaged over all observations with `S = 1`, not only over
the observations whose treatment status changes.

```{r estimate-policy-effect}
policy_est <- counterfactual_policy_effect(
  result = result,
  bootstrap_result = bootstrap_result,
  policy = counterfactual_policy,
  Y = dat$Y,
  X = dat$X,
  D = dat$D,
  level = 0.90
)
```

Because this is a simulation, the true policy effect can also be calculated
using the known conditional mean potential outcomes. For comparison with the
estimator, the true effect is averaged over the same observations with `S = 1`.

```{r compare-policy-effect}
D_counterfactual <- counterfactual_policy(dat$X)

true_factual_mean <- ifelse(
  dat$D == 1,
  dat$mu1,
  dat$mu0
)

true_counterfactual_mean <- ifelse(
  D_counterfactual == 1,
  dat$mu1,
  dat$mu0
)

S <- as.numeric(result$S == 1)

true_policy_effect <- sum(
  S * (true_counterfactual_mean - true_factual_mean)
) / sum(S)

policy_summary <- data.frame(
  quantity = c(
    "True policy effect",
    "Estimated policy effect",
    "90% confidence interval lower",
    "90% confidence interval upper",
    "Number with identified CATE",
    "Expected number affected",
    "Affected share among identified observations"
  ),
  value = c(
    true_policy_effect,
    policy_est$estimate,
    policy_est$conf_low,
    policy_est$conf_high,
    policy_est$identified_n,
    policy_est$num_affected,
    policy_est$affected_share
  )
)

knitr::kable(policy_summary, digits = 4)
```

## Main functions

| Function | Purpose |
|:--|:--|
| `RDD_extrapolate_CV_band()` | Estimate conditional mean potential outcomes, the \(q_0\) and \(q_1\) mappings, and the support indicator \(S\). |
| `RDD_extrapolate_bootstrap()` | Generate multiplier-bootstrap draws for the mappings and imputed potential outcomes. |
| `plot_q_results()` | Plot \(\widehat q_0\), \(\widehat q_1\), and the comparison between \(\widehat q_1\) and \(\widehat q_0^{-1}\). |
| `plot_idf_region()` | Plot the identified and unidentified regions when `X` has exactly two columns. |
| `counterfactual_policy_effect()` | Estimate the effect of a user-specified deterministic or probabilistic counterfactual treatment policy. |

See the function documentation for complete argument definitions and returned
objects:

```r
?RDD_extrapolate_CV_band
?RDD_extrapolate_bootstrap
?plot_q_results
?plot_idf_region
?counterfactual_policy_effect
```

## Reference

The methodology implemented in this package is developed in:

Ben Deaner and Soonwoo Kwon. **“Extrapolation in Regression Discontinuity
Design Using Comonotonicity.”** 2025.
[arXiv:2507.00289](https://arxiv.org/abs/2507.00289).

```bibtex
@article{deanerkwon2025,
  title        = {Extrapolation in Regression Discontinuity Design Using Comonotonicity},
  author       = {Deaner, Ben and Kwon, Soonwoo},
  year         = {2025},
  eprint       = {2507.00289},
  archivePrefix = {arXiv},
  primaryClass = {econ.EM}
}
```
